\documentclass{article}
\usepackage[
    letterpaper,
    top=1in,
    bottom=1in,
    inner=0.75in,
    outer=0.75in,
    % margin=1in,
]{geometry}

%%%% Page Header and Foot %%%%
\usepackage{fancyhdr}
\fancypagestyle{plain}{
\fancyhf{}
\fancyhead[L]{Rutgers University}
\fancyhead[C]{16:332:545 Digital Communications}
\fancyhead[R]{Spring 2024}
\fancyfoot[L]{}
\fancyfoot[C]{\thepage}
\fancyfoot[R]{}
}
\pagestyle{plain}

%%%% Load Format %%%%
\input{../../Formats/Format.tex}
\input{../../Formats/Macro.tex}
\allowdisplaybreaks

%%%% Theorem Style %%%%
\declaretheorem[numbered=no, style=plain]{axiom, lemma}
\declaretheorem[numberwithin=section,style=definition]{definition}
\declaretheorem[sibling=definition]{theorem, corollary, proposition, conjecture}
\declaretheorem[numbered=no,style=remark]{remark, claim}

%%%% Document Information %%%%
\title{Project on OFDM Communications}
\author{Kailong Wang}
\date{\today}

\addbibresource{bib.bib}
% \graphicspath{ {../MyOFDM_1/} }
% \epstopdfsetup{outdir=./}

%%%% Start Document %%%%
\begin{document}
\maketitle

\begin{abstract}
    Orthogonal Frequency Division Multiplexing (OFDM)~\cite{5307460} is a popular modulation scheme in modern wireless communications. It achieves high symbol rates while being robust to frequency-selective fading and intersymbol interference. This project discusses the OFDM technique and evaluates its performance with simulations. The project consists of following parts as required:
    \begin{enumerate}[(i)]
        \item A description of the OFDM modulation and demodulation process.
        \item A discussion of the advantages and disadvantages of OFDM technique.
        \item The usage of pilot symbols for channel estimation in OFDM system.
        \item Numerical results to evaluates the OFDM scheme.
    \end{enumerate}
\end{abstract}

\section{Modulation and Demodulation}
\label{sec:modulation}
\input{01sec1.tex}

\section{Advantages and Disadvantages}
\input{02sec2.tex}

\section{Synchronization and Pilot Symbols Aided Channel Estimation}
\input{03sec3.tex}

\section{Simulation Results}
\input{04sec4.tex}

\section{Conclusion}
In this report, we introduce the OFDM modulation and demodulation process, discuss its mathematical properties, and evaluate its performance with simulations. We show that the OFDM system is robust to frequency-selective fading and intersymbol interference. The simulation results demonstrate that the OFDM system is only subject to flat fading as long as the cyclic prefix length is longer than the channel delay spread. The OFDM system is widely used in modern wireless communications due to its high symbol rates and robustness to channel impairments.

\clearpage
\printbibliography

\end{document}